---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BiSbmSim

<!-- badges: start -->
<!-- badges: end -->

This package simulates data using the bipartite stochastic block model (BiSbm). 


## Installation

You can install the unrelased version of bisbmsim from github with:

``` r
devtools::install_github('tbilab/bisbmsim')
```

```{r setup}
library(bisbmsim)

library(tidyverse)
library(purrr)
```


## Example

Super simple simulation. 

First we need to setup basic parameters corresponding to model
```{r}
N_a <- 125  # Number of nodes of the a type
N_b <- 105  # Number of nodes of the b type
K_a <- 4    # How many blocks of a type nodes are there
K_b <- 3    # How many blocks of the b type nodes are there
```

In the future these variables will be chosen, not generated.
```{r}
b_a <- assign_group_membership(N = N_a, K = K_a) # Block membership for each a node (automated in future)
b_b <- assign_group_membership(N = N_b, K = K_b) # Block membership for each b node
```

### Average number of connections between each pair of a and b groups
```{r}
Lambda <- generate_random_lambda(K_a = K_a, K_b = K_b)
```



Look at what our Lambda looks like real quick.

```{r visualizing_lambda}
ggplot(Lambda, aes(x = a, y = b)) +
  geom_tile(
    aes(fill = avg_num_cons),
    color = 'white'
  ) +
  scale_fill_gradient(low = "white", high = "#56B1F7")
```



## Now we can begin simulation

```{r}
all_node_pairs <- draw_from_model(b_a, b_b, Lambda)
all_node_pairs %>% head()
```


Plot results and compare with the generating lambda matrix. 
```{r simulation_results}
all_node_pairs %>% 
  mutate(num_edges = ifelse(num_edges >= 1, 1, 0)) %>% 
  gather(key = 'type', value = 'connections', num_edges) %>% 
  arrange(a_group, b_group) %>% 
  ggplot(aes(x = reorder(a, a_group), y = reorder(b, b_group))) +
  geom_tile(aes(fill = connections) ) +
  facet_wrap(~type) +
  scale_fill_gradient(low = "white", high = "#56B1F7") + 
  labs(
    x = 'a node',
    y = 'b node'
  ) +
  theme_minimal() +
  theme(
    axis.text = element_blank()
  )
```
